---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# csmp

<!-- badges: start -->
<!-- badges: end -->

The goal of csmp is to assess the climate sensitivity of model performance, as described in Marshall et al. (2021). Please see the manuscript for more details about use and applications. 

## Installation

You can install csmp from [GitHub](https://github.com/) with:

``` {r installation}
# install.packages("devtools") # if devtools is not already installed
devtools::install_github("adrienne-marshall/CSMP", ref = 'main') # ref needed for older versions of devtools()
```

## Examples

We can evaluate the model performance, as represented by an objective function (e.g., RMSE), in different climatic contexts. Each value in the objective function vector represents model performance in the corresponding air temperature and precipitation vectors. In the following example, air temperature, precipitation, and the objective function are all random vectors representing hypothetical annual mean air temperature, annual precipitation, and the model performance in a given year. 

```{r example}
library(csmp)
set.seed(10)
airtemp <- rnorm(20, mean = 0, sd = 1)
precip <- rnorm(20, mean = 500, sd = 100)
objective <- runif(20, min = 0, max = 1)

ans <- csmp(airtemp, precip, objective, return_plot = TRUE)
ans$stats
ans$contour_plot 

```

The output contour plot can be modified with standard ggplot commands: 

```{r modify plot}
library(ggplot2)
ans$contour_plot + 
  labs(fill = "Objective\nfunction",
       x = "Air temperature (°C)",
       y = "Precipitation (mm)")
```

The `new_airtemp` and `new_precip` parameters represent the conditions under which to compare the change in model performance. For example, for evaluation of model performance in a 2 °C warming scenario, `new_airtemp` should be 2 °C warmer than historical average air temperature (which can be specified with the `ref_airtemp` parameter or inferred based on the supplied vector of air temperature observations). `new_airtemp` and `new_precip` can each be vectors of length one (representing unidirectional change in climate) or two (to evaluate, e.g., potential changes due to wetting or drying).

In this example, we modify the new values in air temperature and precipitation. Keep in mind the model will be most reliable if predictions are within the range of observed variability. 

```{r modify example}
# ans <- csmp(airtemp, precip, objective,
#             return_plot = TRUE,
#             new_airtemp = c(-4, 4),
#             new_precip = c(100, 800))
# ans$stats
```

# References

Marshall, A.M., T.E. Link, G.N. Flerchinger, D.J. Nicolsky, M.S. Lucash (2021). Ecohydrologic modeling in deciduous boreal forest: Model evaluation for application in non-stationary climates. _*Submitted to Hydrological Processes*_.


